---
title: 'Script: SpatialACItaly'
author: "Michael Tiefelsdorf"
date: "4/05/2021"
output: 
  html_document:
    toc: true
    toc_float: 
          collapsed: false
    number_sections: true          
    toc_depth: 3  
---
```{r extractCode}
##
## Extract R code from R markdown documents
##
## setwd("E:\\Lectures2021\\GISC7310\\GLSHeteroSpaAC")
## knitr::purl("SpatialACItaly.rmd")
##
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, prompt = TRUE, comment = "R>", 
                      cache = FALSE, fig.width=7.2)
rm(list=ls(all=TRUE))
library(car); library(maptools); library(spdep); library(TexMix); library(spatialreg)
```

# Import data
Import the shape-files and run basic diagnostics.
```{r import}
setwd("E:\\Lectures2021\\GISC7310\\GLSHeteroSpaAC\\ItalyMaps")
##
## Read Poly Shapefiles (readShapePoly in library maptools)
##
getinfo.shape("Provinces.shp")
neig.shp <- rgdal::readOGR(dsn=getwd(), layer="Neighbors", integer64="warn.loss", stringsAsFactors=TRUE)
prov.shp <- rgdal::readOGR(dsn=getwd(), layer="Provinces", integer64="warn.loss", stringsAsFactors=TRUE)
summary(prov.shp)

proj4string(prov.shp)                                    # map projection
prov.centroid <- coordinates(prov.shp)                   # Get province centroids
prov.bbox <- bbox(prov.shp)                              # province bounding box for map region
```

# Qualitative map theme of `REGION`
```{r qualmap, fig.width=7, fig.height=9}
is.factor(prov.shp$REGION)
table(prov.shp$REGION)

plot(neig.shp,axes=T,col=grey(0.9), border="white",      # background: neighboring countries
     xlim=prov.bbox[1,], ylim=prov.bbox[2,])                     
mapColorQual(prov.shp$REGION, prov.shp, map.title="Italy's Regions",
                          legend.title="Region", add.to.map=T)        

```

# Gradient map theme of `TOTFERTRAT`
```{r gradmap, , fig.width=7, fig.height=9}
plot(neig.shp,axes=T,col=grey(0.9),border="white",       # first background (axes=T adds lat/long frame)
     xlim=prov.bbox[1,],ylim=prov.bbox[2,])              # within bounding box
mapColorRamp(prov.shp$TOTFERTRAT, prov.shp, breaks=8,    # second add map
              map.title="Spatial Pattern of Fertility Rate",
              legend.title="Fertility Rate", 
              legend.cex=0.7, add.to.map=T)

```

# Basic regression model
```{r basicreg, fig.width=9, fig.height=9}
## Regression model for fertility including model diagnostics
scatterplotMatrix(~TOTFERTRAT+ILLITERRAT+FEMMARAGE+DIVORCERAT+TELEPERFAM, data=prov.shp,
                   smooth=list(span = 0.35, lty.smooth=1, col.smooth="red", col.var="salmon"),
                   regLine=list(col="green"))

fert.lm <- lm(TOTFERTRAT~ILLITERRAT+FEMMARAGE+DIVORCERAT+TELEPERFAM, data=prov.shp)
summary(fert.lm,corr=F)
vif(fert.lm)

# Perform Residual Diagnostics
influenceIndexPlot(fert.lm, id=list(n=3,labels=prov.shp$PROVNAME))
fertResid <- residuals(fert.lm)
```

# Identify potential outlier
```{r outlier, fig.width=7, fig.height=9}
## Why is Bolzano-Bozen an extreme observation? Shall we delete it?
idx.max <- which.max(abs(fertResid))        # Get index of a record with "outlying" observation 

## Map potential outlier
extremeObs <- rep(0, length(fertResid))
extremeObs[idx.max] <- 1
extremeObs <- factor(extremeObs, labels=c("Population","Outlier"))
table(extremeObs)
plot(neig.shp,axes=T,col=grey(0.9),border="white",                 # background: neighboring countries
     xlim=prov.bbox[1,],ylim=prov.bbox[2,])                     
mapColorQual(extremeObs, prov.shp, map.title="Potential Outlier",
              legend.title="Outliers", add.to.map=T)        

## Inspect outlier
prov.shp@data[idx.max,]                      # List info of record centre is outlier

## Delete outlier or update information
#prov.shp <- prov.shp[-idx.max ]             # Delete extreme observation from shapefile 
prov.shp@data[idx.max, "TOTFERTRAT"] <- 1.2  # Or change its value within the shapefile

## Continue with updated dataset
fert.lm <- lm(TOTFERTRAT~ILLITERRAT+FEMMARAGE+DIVORCERAT+TELEPERFAM,data=prov.shp) # update model
summary(fert.lm)
influenceIndexPlot(fert.lm, id=list(n=3, labels=prov.shp$PROVNAME))
```

# Check for heteroscedasticity
```{r heterocheck}
## Check for heteroscedasticity
fert.fgls <- lmHetero(TOTFERTRAT~ILLITERRAT+FEMMARAGE+DIVORCERAT+TELEPERFAM | log(FEMPOP94),
                      data=prov.shp)
summary(fert.fgls)
```
Note the positive sign of $\gamma_2$ is contradicting to the theory, which implies that larger areas have lower variances. 

This is marginal significant, thus the heteroscedastic model and will be used subsequently for demonstration purposes.

Therefore, the lm-model is updated to a __weighted lm-model__.
```{r wlm}
fert.wlm <- lm(TOTFERTRAT~ILLITERRAT+FEMMARAGE+DIVORCERAT+TELEPERFAM, data=prov.shp,
               weights=fert.fgls$weights)
```

# Bipolar map theme of regression residual
```{r residmap, fig.width=7, fig.height=9}
## Plot Regression Residuals (Bi-polar)
fertResid <- weighted.residuals(fert.wlm)                # Update residuals
hist(fertResid, main="Residuals of Weighted Model")      # Explore distribution to
length(fertResid[fertResid < 0])                         # identify number of pos/neg classes
length(fertResid[fertResid >= 0])

plot(neig.shp,axes=T,col=grey(0.9),border="white",
     xlim=prov.bbox[1,],ylim=prov.bbox[2,])               # first background
mapBiPolar(fertResid, prov.shp,                           # second regression residuals
            neg.breaks=5, pos.breaks=5, break.value=0.0, 
            map.title="Fertility Model Residuals",
            legend.title="Residuals", 
            legend.cex=0.7, add.to.map=T)
```

The map pattern indicates that there are spatial clusters with positive and negative regression residuals. Thus the __spatial independence assumption__ is violated.

# Identify the linkage structure
```{r linkmap, fig.width=7, fig.height=9}
## Plot Augmented Spatial Links among Italian Provinces
## Notes: Shape file has been edited so satellite islands are connected to mainland
## Alternatively spdep::edit.nb function (does not work with RStudio)
prov.link <- poly2nb(prov.shp, queen=F)                          # Generate neighbors links

plot(neig.shp,axes=T,col=grey(0.9),border="white",
     xlim=prov.bbox[1,],ylim=prov.bbox[2,])                      # First background
plot(prov.shp,col="palegreen3" ,border=grey(0.9), axes=T, add=T) # Second plot areas
plot(prov.link,coords=prov.centroid, pch=19, cex=0.1,            # Third plot links focused at centroids
     col="blue", add=T)
title("Augmented Spatial Links among Provinces")                 # Forth add title
box()                                                            # Fifth refresh frame
```

# Perform a test for spatial autocorrelation in the residuals
```{r morantest, fig.width=9, fig.height=9}
prov.linkW <- nb2listw(prov.link, style="W")                   # generated row-sum standardized neighbors list
spOutliers <- moran.plot(weighted.residuals(fert.wlm),         # Moran plot with outlier diagnositics
                         prov.linkW, labels=prov.shp$PROVNAME)          
lm.morantest(fert.wlm, prov.linkW)                             # Test with W-coding scheme
```
Also look up `help(lm.morantest)`. Internally weighted regression residuals are used.

# Maximum likelihood estimation of SAR
Check the online help for the `spatialreg::spautolm( )` function. Note the use of the `weights` option here. When the `weights` option is used `spautolm` requires a full specification of the model rather than just using the `fert.wlm` object.
```{r SARmodel}
fert.SAR <- spautolm(TOTFERTRAT~ILLITERRAT+FEMMARAGE+DIVORCERAT+TELEPERFAM, 
                     data=prov.shp, weights = fert.fgls$weights,
                     na.action="na.omit", listw=prov.linkW, family="SAR")
summary(fert.SAR)
```

The likelihood ratio test of the SAR model is identically to the hand calculated one. Important here is a slight difference in the calculation when heteroscedasticity is incorporated.
```{r SARlikeratio}
## Likelihood Ratio test (identical to LR from spautolm)
#(likeH0 <- logLik(fert.lm))                     # Use for unweighted model
(likeH0 <- fert.fgls$logLikeH1)
(likeH1 <- logLik(fert.SAR))

cat("chi-square value:  ", chi <- -2*(likeH0[1]-likeH1[1]))
cat("error-probability: ", pchisq(chi, df=1, lower.tail=F))
```

Thus the SAR model improves significantly over the heteroscedastic model.

Check whether the residuals are as expected free of spatial autocorrelation. Because the distribution of ML residuals in unknown, Morans's $I$ is evaluated here using a __randomization approach__.
```{r SARresid, fig.width=7, fig.height=9}
## Moran test applying randomization because ML may not be normal distributed
plot(neig.shp,axes=T,col=grey(0.9),border="white",
     xlim=prov.bbox[1,],ylim=prov.bbox[2,])                        # first background
mapBiPolar(residuals(fert.SAR), prov.shp,                          # second regression residuals
            neg.breaks=5, pos.breaks=5, break.value=0.0, 
            map.title="Fertility SAR Model Residuals",
            legend.title="Residuals", legend.cex = 0.6,
            add.to.map=T)

## Evaluate ML residuals for spatial autocorrelation
moran.mc(residuals(fert.SAR), prov.linkW, nsim=9999, alternative="less")  
```

