---
title: "Trend Surface"
author: "Yalin Yang"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: no
    toc_depth: '3'
  html_notebook:
    toc: yes
    toc_float: yes
  word_document:
    toc: no
    toc_depth: '3'
---


## 1 Polynom Correlation

**Correlation between higher order polynoms**

### 1.1 Positive branch

```{r}
x1 <- seq(0,2,by=0.05)
x2 <- x1^2
x3 <- x1^3

plot(x2~x1, type="l", col="green", main="Possitive Branch")
lines(x3~x1, col="red")
```

```{r}
cor(x2,x1)
```
```{r}
cor(x3,x1)
```

### 1.2 Centered around zero

```{r}
x1 <- seq(-1,1,by=0.05) # note shift of scale
x2 <- x1^2
x3 <- x1^3

plot(x2~x1, type="l", ylim=range(c(x2,x3)), col="green", main="Centered around Zero")
lines(x3~x1, col="red")
abline(v=0, h=0, lty=5)
```
```{r}
cor(x2,x1)
```
```{r}
cor(x3,x1)
```

## Trend Surface Model

### 1.1 Transfer Coordinates to a comparable range

* Objective: z-transforms a coordinate (reduce multicollinearity among the polynomial transformed coordinates)
* `x`: vector of coordinates
* `xmean`: centers around xmean
* `xsd`: scale factor as standard deviation

```{r}
zTransCoord <- function(x,xmean,xsd,reverse=F){  
  if (reverse == F) x <- (x-xmean)/xsd
  else x <- x*xsd+xmean
  return(x)
} 

x <- rnorm(5,mean = 2,sd = 5)
x1 <- zTransCoord(x,xmean = 2,xsd = 5,reverse = F)
(demo <- data.frame(orign = x, transfered = x1))

```

### 1.2 Generate formula for polygonial trend-surface model

* Objective: Generate formula for polygonial trend-surface model (multiple order)
* `baseForm`: Basic formula defining the dependent variable with an intercept, i.e., `Z~1` (zero order). Or alternatively with covariables, e.g., `Z~V1+V2+I(V2^2)`(second order)
* `coordForm`: Right-hand formula defining the longitude and latitude coordinate names, i.e., ~X+Y
* `shpDf`:  optional dataframe (perhaps in geo-object) to the variables names in baseForm and coordForm 
* `polyDeg`: Degree of the trendsurface model (minimum is 1)

**Example:   `makeTrendPolyFct(elev~rivDist, ~long+lat, kansas, polyDeg=3)` **

```{r}
makeTrendPolyForm <- function(baseForm=Z~1, coordForm=~X+Y, shpDf=NULL, polyDeg=1){

  ## Basic input check
  if (missing(baseForm) || class(baseForm) != "formula") stop("'baseForm' missing or incorrect")
  if (missing(coordForm) || class(coordForm) != "formula" || length(all.vars(coordForm)) != 2)
    stop("'coordForm' missing or incorrect")
  if (as.integer(polyDeg) < 1L ) stop("'polyDeg' needs to be integer of 1 or above")
  ## shpDf declare check input variables
  if (!is.null(shpDf)){
    if (class(try(model.frame(baseForm, shpDf), T)) != "data.frame") stop("Incorrect variables in 'baseFrom'")
    if (class(try(model.frame(coordForm, shpDf), T)) != "data.frame") stop("Incorrect variables in 'coordFrom'")
  }
  ## Build string with the trend-surface coordinates
  expo <- expand.grid(0L:polyDeg, 0L:polyDeg)
  expo <- expo[rowSums(expo) <= polyDeg, ]    # make sure only terms up to polyDeg enter the formula
  expo <- expo[-1,]                           # exclude x^0*y^0
  expo <- expo[order(rowSums(expo)), ]        # make sure formula is sorted by the degree of the polynomial
  formStr <- "~ ."
  xy <- all.vars(coordForm)  
  for (i in 1: nrow(expo)) {
    formStr <- paste(formStr, " + I(",xy[1],"^",expo[i,1],"*",xy[2],"^",expo[i,2],")", sep="")  
  }  
    ## Merge covariate and polynomial formulas
  polyForm <- update(baseForm,as.formula(formStr))
  
  return(polyForm)
} #end::makeTrendPolyForm
```


### 1.3 Read Map Data

```{r message=FALSE, warning=FALSE}
library(maptools); library(RColorBrewer)

setwd("G:\\UTD_Classes\\2020Spring\\GISC7360_Pattern_Analysis\\Lecture05\\TrendShapeFiles")
landBnd <- rgdal::readOGR(dsn=getwd(), layer="Land", integer64="warn.loss")
proj4string(landBnd)                              # Current system
statPts <- rgdal::readOGR(dsn=getwd(), layer="BritStations", integer64="warn.loss")
proj4string(statPts)
elevGrid <- rgdal::readOGR(dsn=getwd(), layer="Elevation", integer64="warn.loss") 
proj4string(elevGrid)
```

**read the boundary information**

```{r}
( boxPred <- bbox(elevGrid) )
( boxCali <- bbox(statPts) )
```

### 1.4 Map Study Area

```{r}
plot(landBnd,axes=T,col=grey(0.9),border="green",
     xlim=boxPred[1,],ylim=boxPred[2,])
plot(statPts,col="red",add=T)
plot(elevGrid,col="black",pch=".",add=T)
box()
```

**Transform coordinates**

```{r}
yMean <- mean(statPts$LATITUDE1); xMean <- mean(statPts$LONGITUDE1)
ySd <- sd(statPts$LATITUDE1); xSd <- sd(statPts$LONGITUDE1)
```


#### 1.4.1 Test no transformation

```{r}
statPts$caliX <- statPts$LONGITUDE1
statPts$caliY <- statPts$LATITUDE1
elevGrid$caliX <- elevGrid$LONGITUDE1
elevGrid$caliY <- elevGrid$LATITUDE1
```

**1st Order Trendsurface**

```{r}
polyForm1 <- makeTrendPolyForm(SEPTTEMP~ALT, coordForm =~caliX+caliY, shpDf=statPts, polyDeg=1)
polyForm1
```

```{r}
lmTrend1 <- lm(polyForm1, data=statPts)
summary(lmTrend1)
```

**2nd Order TrendSurface**

```{r}
polyForm2 <- makeTrendPolyForm(SEPTTEMP~ALT, ~caliX+caliY, shpDf=statPts, polyDeg=2)
lmTrend2 <- lm(polyForm2, data=statPts)
summary(lmTrend2)
```

**Perform partial F-test: Compare first and second order trendsurface**

```{r}
anova(lmTrend1, lmTrend2)
```

**3rd Order TrendSurface**

```{r}
polyForm3 <- makeTrendPolyForm(SEPTTEMP~ALT, ~caliX+caliY, shpDf=statPts, polyDeg=3)
lmTrend3 <- lm(polyForm3, data=statPts)
summary(lmTrend3)
```

**check multicollinearity**

```{r}
car::vif(lmTrend3)
```

**Perform partial F-test: Compare second and third order trendsurface**

```{r}
anova(lmTrend2, lmTrend3)
```

**Best model is 2nd order**

### 1.5 Map 2nd Order(best model) Trend Surface

```{r}
## Perform Prediction
predTrend2 <- predict(lmTrend2, elevGrid, se.fit = TRUE)
predFit <- predTrend2$fit

n.col <- 9 # number of color
pal <- brewer.pal(n.col,"YlOrRd")
predClass <- classInt::classIntervals(predFit, n.col, style="equal")
plot(predClass, pal=pal)
```

**Map**

```{r fig.height=8, fig.width=13}
predCol <- classInt::findColours(predClass,pal)

plot(elevGrid,axes=T,col=predCol,pch=15,cex=2,
     xlim=boxPred[1,],ylim=boxPred[2,],
     main="Predition: 2nd Order Trendsurface")
plot(landBnd,axes=T,border="grey50",add=T)
plot(statPts,col="blue",add=T)
legend("topright", title = "Temperatur Predictions", 
       legend = leglabs(round(predClass$brks, digits = 0)), fill = pal, 
       bty = "o", ncol = 1)
```

### 1.6 Map 2nd Prediction Uncertainty

** Cannot map residuals because the prediction grid does not have observed temperature data (not all area have observed value)
```{r}
predSe <- predTrend2$se
n.col <- 9
pal <- brewer.pal(n.col,"Reds")
seClass <- classInt::classIntervals(predSe, n.col, style="equal")
plot(seClass,pal=pal)
seCol <- classInt::findColours(seClass,pal)
```

```{r fig.height=8, fig.width=13}
plot(elevGrid,axes=T,col=seCol,pch=15,cex=2,
     xlim=boxPred[1,],ylim=boxPred[2,],
     main="Standard Error: 2nd Order Trendsurface")
plot(landBnd,axes=T,border="green",add=T)
plot(statPts,col="blue",add=T)
legend("topright", title = "Prediction Std.-Erros", 
       legend = leglabs(round(seClass$brks, digits = 0)), fill = pal, 
       bty = "o", ncol = 1)
```

