---
title: 'Script: SpatialACItaly'
author: "Michael Tiefelsdorf"
date: "3/29/2020"
output: 
  html_document:
    toc: true
    toc_float: 
          collapsed: false
    number_sections: true          
    toc_depth: 3  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include=TRUE, highlight=TRUE, tidy=TRUE, fig.width=6)

rm(list=ls(all=TRUE))
library(car); library(maptools); library(spdep); library(TexMix)
```

# Import data
Import the shape-files and run basic diagnostics.
```{r import}
setwd("G:\\UTD_Classes\\2020Spring\\GISC7360_Pattern_Analysis\\Lecture06\\Italy")
##
## Read Poly Shapefiles (readShapePoly in library maptools)
##
getinfo.shape("Provinces.shp")
neig.shp <- rgdal::readOGR(dsn=getwd(), layer="Neighbors", integer64="warn.loss")
prov.shp <- rgdal::readOGR(dsn=getwd(), layer="Provinces", integer64="warn.loss")
summary(prov.shp)

proj4string(prov.shp)                                    # map projection
prov.centroid <- coordinates(prov.shp)                   # Get province centroids
prov.bbox <- bbox(prov.shp)                              # province bounding box for map region
```

# Gradient map of `TOTFERTRAT` and bi-polar map of `lm(TOTFERTRAT~1)`
Gradient map:
```{r gradmap}
plot(neig.shp,axes=T,col=grey(0.9),border="white",       # first background (axes=T adds lat/long frame)
     xlim=prov.bbox[1,],ylim=prov.bbox[2,])              # within bounding box
mapColorRamp(prov.shp$TOTFERTRAT, prov.shp, breaks=8,    # second add map
              map.title="Spatial Pattern of Fertility Rate",
              legend.title="Fertility Rate", 
              legend.cex=0.7, add.to.map=T)

```

Variation around the mean:
```{r}
varMean <- residuals(lm(TOTFERTRAT~1, data=prov.shp))
plot(neig.shp,axes=T,col=grey(0.9),border="white",
     xlim=prov.bbox[1,],ylim=prov.bbox[2,])               # first background
mapBiPolar(varMean, prov.shp,                             # second regression residuals
            neg.breaks=5, pos.breaks=5, break.value=0.0, 
            map.title="Fertility Variation Around the Mean",
            legend.title="Residuals", 
            legend.cex=0.7, add.to.map=T)
```

# Basic regression model
```{r basicreg}
## Regression model for fertility including model diagnostics
scatterplotMatrix(~TOTFERTRAT+ILLITERRAT+FEMMARAGE+DIVORCERAT+TELEPERFAM, data=prov.shp,
                   smooth=list(span = 0.35, lty.smooth=1, col.smooth="red", col.var="salmon"),
                   regLine=list(col="green"))

fert.lm <- lm(TOTFERTRAT~ILLITERRAT+FEMMARAGE+DIVORCERAT+TELEPERFAM, data=prov.shp)
summary(fert.lm,corr=F)
vif(fert.lm)

# Perform Residual Diagnostics
influenceIndexPlot(fert.lm, id=list(n=3,labels=prov.shp$PROVNAME))
fertResid <- residuals(fert.lm)
```

# Identify potential outlier
```{r outlier}
## Why is Bolzano-Bozen an extreme observation? Shall we delete it?
idx.max <- which.max(abs(fertResid))        # Get index of a record with "outlying" observation 

## Map potential outlier
extremeObs <- rep(0, length(fertResid))
extremeObs[idx.max] <- 1
extremeObs <- factor(extremeObs, labels=c("Population","Outlier"))
table(extremeObs)
plot(neig.shp,axes=T,col=grey(0.9),border="white",                 # background: neighboring countries
     xlim=prov.bbox[1,],ylim=prov.bbox[2,])                     
mapColorQual(extremeObs, prov.shp, map.title="Potential Outlier",
              legend.title="Outliers", add.to.map=T)        

## Inspect outlier
prov.shp@data[idx.max,]                      # List info of record centre is outlier

## Delete outlier or update information
#prov.shp <- prov.shp[-idx.max ]             # Delete extreme observation from shapefile 
prov.shp@data[idx.max, "TOTFERTRAT"] <- 1.2  # Or change its value within the shapefile

## Continue with updated dataset
fert.lm <- lm(TOTFERTRAT~ILLITERRAT+FEMMARAGE+DIVORCERAT+TELEPERFAM,data=prov.shp) # update model
summary(fert.lm)
influenceIndexPlot(fert.lm, id=list(n=3, labels=prov.shp$PROVNAME))
```

# Check for heteroscedasticity
```{r heterocheck}
## Check for heteroscedasticity
fert.fgls <- lmHetero(TOTFERTRAT~ILLITERRAT+FEMMARAGE+DIVORCERAT+TELEPERFAM | log(FEMPOP94),
                      data=prov.shp)
summary(fert.fgls)
```
Note the possitive sign of $\gamma_2$ is contradicting theory. This is marginal significant of the heteroscedastic model and will be used subsequently for demonstration purposes.

Therefore, the lm-model is updated to a weighted lm-model.
```{r wlm}
fert.wlm <- lm(TOTFERTRAT~ILLITERRAT+FEMMARAGE+DIVORCERAT+TELEPERFAM, data=prov.shp,
               weights=fert.fgls$weights)
```

# Bipolar map theme of regression residual
```{r residmap}
## Plot Regression Residuals (Bi-polar)
fertResid <- weighted.residuals(fert.wlm)                # Update residuals
hist(fertResid, main="Residuals of Weighted Model")      # Explore distribution to
length(fertResid[fertResid < 0])                         # identify number of pos/neg classes
length(fertResid[fertResid >= 0])

plot(neig.shp,axes=T,col=grey(0.9),border="white",
     xlim=prov.bbox[1,],ylim=prov.bbox[2,])               # first background
mapBiPolar(fertResid, prov.shp,                           # second regression residuals
            neg.breaks=5, pos.breaks=4, break.value=0.0, 
            map.title="Fertility Model Residuals",
            legend.title="Residuals", 
            legend.cex=0.7, add.to.map=T)
```

The map pattern indicates that there are spatial clusters with positive and negative regression residuals. Thus the independence assumption is violated.

# Identify the linkage structure
```{r linkmap}
## Plot Augmented Spatial Links among Italian Provinces
## Notes: Shape file has been edited so satellite islands are connected to mainland
## Alternatively spdep::edit.nb function (does not work with RStudio)
prov.link <- poly2nb(prov.shp, queen=F)                          # Generate neighbors links

plot(neig.shp,axes=T,col=grey(0.9),border="white",
     xlim=prov.bbox[1,],ylim=prov.bbox[2,])                      # First background
plot(prov.shp,col="palegreen3" ,border=grey(0.9), axes=T, add=T) # Second plot areas
plot(prov.link,coords=prov.centroid, pch=19, cex=0.1,            # Third plot links focused at centroids
     col="blue", add=T)
title("Augmented Spatial Links among Provinces")                 # Forth add title
box()                                                            # Fifth refresh frame
```

Explore the structure of `prov.link`:
```{r strucLink}
summary(prov.link)
str(prov.link)
```

# Perform a test for spatial autocorrelation in the residuals
```{r morantest}
prov.linkW <- nb2listw(prov.link, style="W")                   # generated row-sum standardized neighbors list
spOutliers <- moran.plot(weighted.residuals(fert.wlm),         # Moran plot with outlier diagnositics
                         prov.linkW, labels=prov.shp$PROVNAME)          
lm.morantest(fert.wlm, prov.linkW)                             # Test with W-coding scheme
```
Also look up `help(lm.morantest)`. Internally weighted regression residuals are used.

# Evaluate Correlogram (correlation at higher order spatial lags)
```{r correl}
plot(sp.correlogram(prov.link, weighted.residuals(fert.wlm), order=8, method="I", style = "W"))
```

From spatial lag 2 onwards Moran's I is no longer significant.

# Test for Spatial Autocorrelation Using Randomization
In case the assumption that the residuals are normally distributed no longer holds, the distribution of Morans's $I$ can evaluated using a randomization approach.
```{r SARresid}
moran.mc(weighted.residuals(fert.wlm), prov.linkW, nsim=9999) 
help(moran.mc, package="spdep")
```

A few examples of random pattern:
```{r randomPat}
##
## Evaluation by randomization
##
fertResid <- weighted.residuals(fert.wlm)

## Original values
mi <- lm.morantest(lm(fertResid~1), prov.linkW)           # Test with W-coding scheme
plot(neig.shp,axes=T,col=grey(0.9),border="white",
     xlim=prov.bbox[1,],ylim=prov.bbox[2,])               # first background
mapBiPolar(fertResid, prov.shp,                           # second regression residuals
           neg.breaks=5, pos.breaks=4, break.value=0.0, 
           map.title="",
           legend.title="Residuals", 
           legend.cex=0.7, add.to.map=T)
title(main=bquote(paste("Observed Fertility Model Residuals with ", I==.(round(mi$estimate[[1]],3)))),
      cex.main=1.2)

## 1st Randomization
fertResid <- sample(fertResid, size=length(fertResid))
mi <- lm.morantest(lm(fertResid~1), prov.linkW)           # Test with W-coding scheme
plot(neig.shp,axes=T,col=grey(0.9),border="white",
     xlim=prov.bbox[1,],ylim=prov.bbox[2,])               # first background
mapBiPolar(fertResid, prov.shp,                           # second regression residuals
           neg.breaks=5, pos.breaks=4, break.value=0.0, 
           map.title="",
           legend.title="Residuals", 
           legend.cex=0.7, add.to.map=T)
title(main=bquote(paste("1st Randomized Fertility Model Residuals with ", I==.(round(mi$estimate[[1]],3)))),
      cex.main=1.2)

## 2nd Randomization
fertResid <- sample(fertResid, size=length(fertResid))
mi <- lm.morantest(lm(fertResid~1), prov.linkW)           # Test with W-coding scheme
plot(neig.shp,axes=T,col=grey(0.9),border="white",
     xlim=prov.bbox[1,],ylim=prov.bbox[2,])               # first background
mapBiPolar(fertResid, prov.shp,                           # second regression residuals
           neg.breaks=5, pos.breaks=4, break.value=0.0, 
           map.title="",
           legend.title="Residuals", 
           legend.cex=0.7, add.to.map=T)
title(main=bquote(paste("2st Randomized Fertility Model Residuals with ", I==.(round(mi$estimate[[1]],3)))),
      cex.main=1.2)

```

# Eigenvalues and Vectors of $M \cdot (V+V^T)/2 \cdot M$
The two most extreme spatial patterns are shown below

```{r}
V <- spdep::nb2mat(prov.link, style="W", zero.policy = TRUE)
X <- model.matrix(fert.wlm)
n <- nrow(X)
M <- diag(1,nrow=n) - X %*% solve(t(X)%*%X) %*% t(X)
MVM <- M %*% ((V+t(V))*0.5) %*% M
eig <- eigen(MVM, symmetric = T)

# First eigenvector
plot(neig.shp,axes=T,col=grey(0.9),border="white",
     xlim=prov.bbox[1,],ylim=prov.bbox[2,])               # first background
mapBiPolar(eig$vectors[,1], prov.shp,                
           neg.breaks=5, pos.breaks=4, break.value=0.0, 
           map.title="",
           legend.title="1st Eigenvector", 
           legend.cex=1, add.to.map=T)
title(main=bquote(paste("1st Eigenvector with ", I==.(round(eig$values[1],3)))),
      cex.main=1.5)

# Last eigenvector
plot(neig.shp,axes=T,col=grey(0.9),border="white",
     xlim=prov.bbox[1,],ylim=prov.bbox[2,])               # first background
mapBiPolar(eig$vectors[,n], prov.shp,                
           neg.breaks=5, pos.breaks=4, break.value=0.0, 
           map.title="",
           legend.title="95th Eigenvector", 
           legend.cex=1, add.to.map=T)
title(main=bquote(paste("95th Eigenvector with ", I==.(round(eig$values[n],3)))),
      cex.main=1.5)
```

